White Dwarf Mergers on Adaptive Meshes\\ II. Nuclear Reactions

### 2018-01-23 papers/wdmerger_II/paper.tex
Simplify the nuclear network section

White dwarfs are primarily composed of the $\alpha$-chain particles ${}^4$He,
${}^{12}$C, ${}^{16}$O, ${}^{20}$Ne, and ${}^{24}$Mg. Therefore the core of
any network appropriate for modeling nuclear burning in white dwarfs will be
these alpha chain nuclides, with the idea being that links up the $\alpha$-chain
will eventually get us to ${}^{56}$Ni, the nuclide responsible for the
energy output of Type Ia supernovae. In this paper we use the \aproxthirteen\
network\ \citep{timmes:1999,timmes:2000}. This includes all thirteen $\alpha$-chain
particles between ${}^4$He and ${}^{56}$Ni. This network was used by \citet{hawley:2012}
and \citet{raskin:2010}. \citet{loren-aguilar:2010} and \citet{garcia-senz:2013} used a
very similar network that additionally included ${}^{60}$Zn. This network has been
ported into a form that is consistent with the \amrex\ codes, in the freely available
\microphysics\ code repository\footnote{\microphysics\ can be obtained at
\url{https://github.com/starkiller-astro/Microphysics}.},
a collection of microphysical routines that are designed to be used in our
hydrodynamics codes. In our \microphysics\ repository we have also ported \aproxnineteen\
\citep{timmes:1999}, which was used in a few previous collision papers, and \aproxtwentyone.
The isotopes added by these networks relative to \aproxthirteen\ are not particularly relevant
to the problem of head-on white dwarf collisions, so we do not discuss results with these networks
here (though we have verified for this problem that these larger networks do not significantly change
the answer). We also make available \isoseven\ \citep{timmes:2000}, which skips the nuclides between
silicon and nickel, and was used by \citet{rosswog:2009}. \isoseven\ gives significantly different
quantitative results than \aproxthirteen\ for this problem, due to the inapplicable assumption of
an equilibrium abundance of the unused isotopes, but the results are still qualitatively similar.

@@

The simulations shown in this work all use the self-heating mode. The hydrostatic
and hybrid burns are only a slight variation on the self-heating burn in practice
for this problem.
The reason that the hydrostatic burn cannot substantially change the outcome is
that even though the temperature is fixed during the burn, it is \textit{not}
constant in the rest of the hydrodynamics step. And in the hydrodynamics step,
the \castro\ algorithm performs multiple EOS calls to ensure that the temperature
is synchronized with the internal energy at various points in the update. So in
practice there is a temperature change due to the burn for our problem. This seems
inevitable for any hydro code unless the hydrodynamics update includes an explicit
equation for $T$ that is independent of what is happening for the internal energy.
The hydrostatic burn does slightly limit the energy release during any given burn,
but summed up over the whole simulation this is not enough to qualitatively affect
the outcome. The hybrid burn does not affect the simulation outcome at all relative
to the hydrostatic burn. The difference between them only applies for zones that have
a net negative energy release from the hydrostatic burn, which occurs only after
we have burned to NSE. The hybrid burn was mainly implemented by \citet{raskin:2010}
for making it easier to numerically complete a burn, rather than because of the
effects it has on the outcome of the simulation. The suppressed burn has separate
issues which we defer to \autoref{sec:unstable_burning}

>>>>

In this paper we use the \aproxthirteen\ network\ \citep{timmes:1999,timmes:2000}.
This includes all thirteen $\alpha$-chain particles between ${}^4$He and ${}^{56}$Ni.
These $\alpha$-chain particles capture the most important pathway connecting a WD
and its thermonuclear detonation. This network was used by \citet{hawley:2012}
and \citet{raskin:2010}. \citet{loren-aguilar:2010} and \citet{garcia-senz:2013} used a
very similar network that additionally included ${}^{60}$Zn. The \aproxthirteen\ network has been
ported into a form that is consistent with the \amrex\ codes in the freely available
\microphysics\ code repository\footnote{\microphysics\ can be obtained at
\url{https://github.com/starkiller-astro/Microphysics}.},
a collection of microphysical routines that are designed to be used in our
hydrodynamics codes. In our \microphysics\ repository we have also ported \aproxnineteen\
\citep{timmes:1999}, which was used in a few previous collision papers, and \aproxtwentyone.
The isotopes added by these networks relative to \aproxthirteen\ are not particularly relevant
to the problem of head-on white dwarf collisions, so we do not discuss results with these networks
here (though we have verified for this problem that these larger networks do not significantly change
the answer). We also make available \isoseven\ \citep{timmes:2000}, which skips the nuclides between
silicon and nickel, and which was used by \citet{rosswog:2009}. \isoseven\ gives significantly different
quantitative results than \aproxthirteen\ for this problem, due to its inapplicable assumption of
an equilibrium abundance of the unsimulated isotopes, but the results are still qualitatively similar.

@@

The simulations shown in this work all use the self-heating mode. The hydrostatic
and hybrid burns are only a slight variation on the self-heating burn in practice
for this problem.
The reason that the hydrostatic burn cannot substantially change the outcome is
that even though the temperature is fixed during the burn, it is \textit{not}
constant in the rest of the hydrodynamics step. And in the hydrodynamics step,
the \castro\ algorithm performs multiple EOS calls to ensure that the temperature
is synchronized with the internal energy at various points in the update. So in
practice there is a temperature change due to the burn for our problem. This seems
inevitable for any hydro code unless the hydrodynamics update includes an explicit
equation for $T$ that is independent of what is happening for the internal energy.
The hydrostatic burn does slightly limit the energy release during any given burn,
but summed up over the whole simulation this is not enough to qualitatively affect
the outcome. The hybrid burn does not affect the simulation outcome at all relative
to the hydrostatic burn. The difference between them only applies for zones that have
a net negative energy release from the hydrostatic burn, which occurs only after
we have burned to NSE. The hybrid burn was mainly implemented by \citet{raskin:2010}
for making it easier to numerically complete a burn, rather than because of the
effects it has on the outcome of the simulation. The suppressed burn has separate
issues which we defer to \autoref{sec:unstable_burning}.
